Manuscript draft

### 2019-06-03 manuscript/manuscript draft.Rmd
updated manuscript draft

---
title: "Manuscript draft"
output: html_document
bibliography: library.bib
---

@@

Globally, forests play an important role in the carbon cycle. Old growth forests alone are estimated to sequester up to 1.4 GtCyr superscript^-1^ [@Malhi1999], making them a significant carbon sink, and am important component of global carbon dioxide budgets [@Luyssaert2008]. Understanding the patterns and mechanisms of forest carbon cycling on a global scale is therefore of high importance in understanding the potential role of forests in mitigating climate change.

It is understood that the productivity of forests shows variation with latitude, with a general trend of increasing productivity towards the tropics, however the exact nature of this pattern and the mechanisms that drive it are poorly understood. Climate can be an important driver of productivity, particularly on a global scale; achieving a better understanding of the climatic controls on forest productivity will help to illuminate potential responses and feedbacks of forests to climate change.

@@

Current knowledge indicates that these different components of productivity show different latitudinal patterns. Gross primary productivity is generally thought to be highest in the tropics. Modelling of global terrestrial ecosystem GPP though upscaling and calibration of eddy flux measurements indicates a peak in GPP in the tropics, with the highest levels in tropical forests [@Beer2010; @Jung2011]. This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and mean annual temperature (MAT) and mean annual precipitation (MAP) [@Luyssaert2007], with the highest GPP values reported in tropical forests. The influence of latitude on global patterns of NPP is less clear than that of GPP. Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics. However, other studies have found the highest values of NPP in temperate forests [@Luyssaert2007; @Huston2009]. Because of the challenges in accurately obtaining unbiased measures of belowground productivity, many studies focus on ANPP in preference to measures of NPP. Studies on global patterns of ANPP are equally inconclusive: Gillman et al. (2015) found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands. These findings were echoed in other studies, which have found weak or no relationships between ANPP and latitude [@Huston2009].

>>>>

---
title: "Manuscript draft"
output:
  pdf_document:
    citation_package: natbib
bibliography: library.bib
csl: apa.csl
---

@@

Globally, forests play an important role in the carbon cycle. Old growth forests alone are estimated to sequester up to 1.4 GtCyr ^-1^ [@Malhi1999], making them a significant carbon sink, and an important component of global carbon dioxide budgets [@Luyssaert2008]. Understanding the patterns and mechanisms of forest carbon cycling on a global scale is therefore of high importance in understanding the potential role of forests in mitigating climate change.

It is understood that the productivity of forests shows variation with latitude, with a general trend of increasing productivity towards the tropics; however the exact nature of this pattern and the mechanisms that drive it are poorly understood. Climate can be an important driver of productivity, particularly on a global scale; achieving a better understanding of the climatic controls on forest productivity will help to illuminate potential responses and feedbacks of forests to climate change.

@@

Current knowledge indicates that these different components of productivity show different latitudinal patterns. Gross primary productivity is generally thought to be highest in the tropics. Modelling of global terrestrial ecosystem GPP though upscaling and calibration of eddy flux measurements indicates a peak in GPP in the tropics, with the highest levels in tropical forests [@Beer2010; @Jung2011]. This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and mean annual temperature (MAT) and mean annual precipitation (MAP) [@Luyssaert2007], with the highest GPP values reported in tropical forests. The influence of latitude on global patterns of NPP is less clear than that of GPP. Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics. However, other studies have found the highest values of NPP in temperate forests [@Luyssaert2007; @Huston2009]. Because of the challenges in accurately obtaining unbiased measures of belowground productivity, many studies focus on ANPP in preference to measures of NPP. Studies on global patterns of ANPP are equally inconclusive: Gillman et al. [-@Gillman2015] found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands. These findings were echoed in other studies, which have found weak or no relationships between ANPP and latitude [@Huston2009].
